Supplement to ``{\mytitle}''

### 2022-11-21 ms.Rnw
Improved replication of model 3

# Create Lee et al (2020a) version of model 3 on sub-data
lee3_sub <- haiti3_lee_et_al()

# This model is only fit for a subset of the data, so we will set data
# that are not part of this subset as NA. This will allow us to estimate
# the likelihood for only the values that were considered during the
# parameter estimation phase.
set_as_na <- lee3_sub@times < lubridate::decimal_date(as.Date('2017-06-10'))
temp_ouest_cases <- lee3_sub@data['casesOuest', ]
temp_ouest_cases[set_as_na] <- NA_real_
lee3_sub@data['casesOuest', ] <- temp_ouest_cases

lee3_all <- haiti3_lee_et_al(start_time = "2010-10-23")

@@

      # Save how long it takes for sub-data
      lee3_results[["sub_pf_time"]] <- system.time({

        # Save all likelihood evaluations
        lee3_results[["sub_ll_evals"]] <- foreach(
          i = 1:lee3_NREP, .combine = c
        ) %dopar% {
          logLik(pfilter(lee3_sub, Np = lee3_NP))  # Get likelihood evaluations
        }
      })

@@

      # Save how long it takes for all data
      lee3_results[["all_pf_time"]] <- system.time({

        # Save all likelihood evaluations
        lee3_results[["all_ll_evals"]] <- foreach(
          i = 1:lee3_NREP, .combine = c
        ) %dopar% {
          logLik(pfilter(lee3_all, Np = lee3_NP))  # Get likelihood evaluations
        }
      })

      lee3_results[["n_params"]] <- 29

@@

mod3_lee_coupled_aic <- 2 * lee3_results$n_params - 2 * max(lee3_results$all_ll_evals)

mod3_lee_coupled_partial_aic <- 2 * lee3_results$n_params - 2 * max(lee3_results$sub_ll_evals)

@@

\begin{tabular}{|c|c|c|c|}
\hline
 & \thead{Model~1} & \thead{Model~2} & \thead{Model~3}
\\
\hline
\hline
\multirow{2}{*}{Log-likelihood} &
  $\Sexpr{myround(mod1_ll, 1)}$ &
  $\Sexpr{myround(mod2_ll, 1)}$ &
  $\Sexpr{myround(mod3_ll, 1)}$ \\
    & ($\Sexpr{myround(mod1_lee_ll, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(h2_fit$leeFit_ll, 1)}$) &
  ($\Sexpr{myround(max(lee3_results[["all_ll_evals"]]), 1)}$)\footnotemark[2]
\\
\hline
Number of &
  $\Sexpr{as.character(mod1_n_params)}$ &
  $\Sexpr{as.character(mod2_n_params)}$ &
  $\Sexpr{as.character(mod3_n_params)}$ \\
 Fit Parameters &
 (\Sexpr{as.character(mod1_lee_n_params)}) &
 ($\Sexpr{as.character(h2_fit$leeFit_n_params)}$) &
 ($\Sexpr{as.character(lee3_results$n_params, 1)}$)
\\
\hline
\multirow{2}{*}{AIC} &
  $\Sexpr{myround(mod1_aic, 1)}$ &
  $\Sexpr{myround(mod2_aic, 1)}$ &
  $\Sexpr{myround(mod3_aic, 1)}$ \\
  & ($\Sexpr{myround(mod1_lee_aic, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(mod2_lee_aic, 1)}$) &
  ($\Sexpr{myround(mod3_lee_coupled_aic, 1)}$)\footnotemark[2]
\\
\hline
\thead{Log-ARMA(2,1) \\ Log-likelihood} &
  $\Sexpr{myround(ARMA_benchmarks$m1, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m2, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m3, 1)}$
\\
\hline
\end{tabular}
\caption{\label{tab:likes}Log-likelihood values for each models compared to their ARMA benchmarks.
Values in parenthesis are corresponding values using \citet{lee20} parameter estimates.  $\textsuperscript{1}$The reported likelihood is an upper bound of the likelihood of the \citet{lee20} model. $\textsuperscript{2}$\cite{lee20} fit Model~3 to a subset of the data (March 2014 onward, excluding data from Ouest in 2015-2016).
On this subset, their model has a likelihood of $\Sexpr{myround(max(lee3_results$sub_ll_evals), 1)}$.
On this same subset, our model has a likelihood of $\Sexpr{myround(mod3_eval_res$subset_ll, 1)}$.
See Sec.~S5 of the supplement material for more details on estimating the likelihood of the \cite{lee20} models.
}
\end{table}

>>>>

# Create Lee et al (2020a) version of model 3 on sub-data
lee3_sub <- lee3_spatPomp()

# This model is only fit for a subset of the data, so we will set data
# that are not part of this subset as NA. This will allow us to estimate
# the likelihood for only the values that were considered during the
# parameter estimation phase.
# set_as_na <- lee3_sub@times < lubridate::decimal_date(as.Date('2017-06-10'))
# temp_ouest_cases <- lee3_sub@data['casesOuest', ]
# temp_ouest_cases[set_as_na] <- NA_real_
# lee3_sub@data['casesOuest', ] <- temp_ouest_cases

lee3_all <- lee3_spatPomp(start_date = "2010-10-23")

@@

      # Save how long it takes for sub-data
      lee3_results[["sub_bpf_time"]] <- system.time({

        lee3_sub_ll_evals <- foreach(
          i = 1:lee3_NREP, .combine = rbind
        ) %dopar% {

          liks <- numeric(10L)

          lee_bpf_out <- bpfilter(lee3_sub, block_size = 1, Np = lee3_NP)
          Ouest_ind <- which(lee_bpf_out@unit_names == "Ouest")

          for (u in 1:10) {
            if (u == Ouest_ind) {
              liks[u] <- sum(lee_bpf_out@block.cond.loglik[u, lee3_sub@times >= lubridate::decimal_date(as.Date('2017-06-10'))])
            } else {
              liks[u] <- sum(lee_bpf_out@block.cond.loglik[u, ])
            }
          }

          liks
        }
      })

      # Save all likelihood evaluations
      lee3_results[["sub_ll_eval"]] <- logmeanexp(rowSums(lee3_sub_ll_evals)[!is.na(rowSums(lee3_sub_ll_evals))])

@@

      # Save how long it takes for all data
      lee3_results[["all_pf_time"]] <- system.time({
        # Save all likelihood evaluations

        lee3_all_ll_evals <- foreach(
          i = 1:lee3_NREP, .combine = c
        ) %dopar% {
          logLik(bpfilter(lee3_all, block_size = 1, Np = lee3_NP))  # Get likelihood evaluations
        }

        lee3_results[["all_ll_eval"]] <- logmeanexp(lee3_all_ll_evals[!is.na(lee3_all_ll_evals)])

      })

      lee3_results[["n_params"]] <- 29  # TODO: Check to make sure this is the correct number.

@@

mod3_lee_coupled_aic <- 2 * lee3_results$n_params - 2 * lee3_results$all_ll_eval

mod3_lee_coupled_partial_aic <- 2 * lee3_results$n_params - 2 * lee3_results$sub_ll_eval

@@

\begin{tabular}{|c|c|c|c|}
\hline
 & \thead{Model~1} & \thead{Model~2} & \thead{Model~3}
\\
\hline
\hline
\multirow{2}{*}{Log-likelihood} &
  $\Sexpr{myround(mod1_ll, 1)}$ &
  $\Sexpr{myround(mod2_ll, 1)}$ &
  $\Sexpr{myround(mod3_ll, 1)}$ \\
    & ($\Sexpr{myround(mod1_lee_ll, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(h2_fit$leeFit_ll, 1)}$) &
  ($\Sexpr{myround(lee3_results[["all_ll_eval"]], 1)}$)\footnotemark[2]
\\
\hline
Number of &
  $\Sexpr{as.character(mod1_n_params)}$ &
  $\Sexpr{as.character(mod2_n_params)}$ &
  $\Sexpr{as.character(mod3_n_params)}$ \\
 Fit Parameters &
 (\Sexpr{as.character(mod1_lee_n_params)}) &
 ($\Sexpr{as.character(h2_fit$leeFit_n_params)}$) &
 ($\Sexpr{as.character(lee3_results$n_params, 1)}$)
\\
\hline
\multirow{2}{*}{AIC} &
  $\Sexpr{myround(mod1_aic, 1)}$ &
  $\Sexpr{myround(mod2_aic, 1)}$ &
  $\Sexpr{myround(mod3_aic, 1)}$ \\
  & ($\Sexpr{myround(mod1_lee_aic, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(mod2_lee_aic, 1)}$) &
  ($\Sexpr{myround(mod3_lee_coupled_aic, 1)}$)\footnotemark[2]
\\
\hline
\thead{Log-ARMA(2,1) \\ Log-likelihood} &
  $\Sexpr{myround(ARMA_benchmarks$m1, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m2, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m3, 1)}$
\\
\hline
\end{tabular}
\caption{\label{tab:likes}Log-likelihood values for each models compared to their ARMA benchmarks.
Values in parenthesis are corresponding values using \citet{lee20} parameter estimates.  $\textsuperscript{1}$The reported likelihood is an upper bound of the likelihood of the \citet{lee20} model. $\textsuperscript{2}$\cite{lee20} fit Model~3 to a subset of the data (March 2014 onward, excluding data from Ouest in 2015-2016).
On this subset, their model has a likelihood of $\Sexpr{myround(lee3_results$sub_ll_eval, 1)}$.
On this same subset, our model has a likelihood of $\Sexpr{myround(mod3_eval_res$subset_ll, 1)}$.
See Sec.~S5 of the supplement material for more details on estimating the likelihood of the \cite{lee20} models.
}
\end{table}

### 2022-11-21 si/si.Rnw
Improved replication of model 3

<<Model3Replication, echo=FALSE, cache=TRUE>>=
NSIM <- switch(run_level, 50, 100, 1000)

@@

lee3 <- haiti3_lee_et_al()
lee3_sims <- simulate(lee3, nsim = NSIM, seed = 321, format = 'data.frame')

@@

quants <- lee3_sims %>%
  dplyr::select(-CasesAll) %>%
  tidyr::pivot_longer(
    data = .,
    cols = starts_with("cases"),
    names_prefix = "cases",
    names_to = "dep",
    values_to = "cases"
  ) %>%
  dplyr::group_by(dep, time) %>%
  dplyr::summarise(
    q05 = quantile(cases, 0.025, na.rm = T),
    mean = mean(cases, na.rm = T),
    q50 = quantile(cases, 0.5, na.rm = T),
    q95 = quantile(cases, 0.975, na.rm = T)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    date = lubridate::date_decimal(time),
    date = as.Date(lubridate::round_date(date))
  )

>>>>

<<Model3Replication, echo=FALSE, cache=TRUE>>=
NSIM <- switch(run_level, 50, 200, 500)

@@

lee3 <- lee3_spatPomp()
lee3_sims <- simulate(lee3, nsim = NSIM, seed = 321, format = 'data.frame')

@@

quants <- lee3_sims %>%
  dplyr::group_by(unitname, time) %>%
  dplyr::summarise(
    q05 = quantile(cases, 0.025, na.rm = T),
    mean = mean(cases, na.rm = T),
    q50 = quantile(cases, 0.5, na.rm = T),
    q95 = quantile(cases, 0.975, na.rm = T)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    date = lubridate::date_decimal(time),
    date = as.Date(lubridate::round_date(date))
  ) %>%
  dplyr::rename(dep = unitname)
